AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-07 Paper/main.tex
10-7 paper

 Specifically, we use a vector $A_i$ to build the first-order preference at position $i$ within sgRNA sequence, then
y=\sum_i^lW_iA_i\cdot X_i\label{eq:04}
 where $W$ denotes a non-negative weight vector, $W_i$ denotes the weight of the $i$-th position, $X_i$ denotes the embedding vector of the nucleotide at $i$-th position.
 We linearly combine the first-order preference vector of base sequence as following
\tilde{A}=BA\label{eq:05}
 where the weight matrix $B$ is learned through attention mechanism, and we define $A$ as the first-order original preference matrix, 
 $\tilde{A}$ as the first-order combine preference matrix (or just first-order preference) to differentiate them.
 In other words, we build a additive model for the first-order preference, 
 which means that the first-order preference of the $i$-th position $\tilde{A}_i$ can be calculated as
\tilde{A}_i=\sum_j^lB_{ij}A_j\label{eq:06}
 Assume that in Equation~(\ref{eq:06}) the first-order original preference at each position is relatively independent and weight matrix B is independent of any first-order original preference, 
 then the total differential of the first-order preference at $i$-th position could be written as
d\tilde{A}_i=\sum_j^lB_{ij}EdA_j\label{eq:07}

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{category.png}}
    \caption{Two categories of the deep learning models used in sgRNA related task. 
    (a) Model work in spatial domain. 
    In spatial domain, the base sequence is encoded into a binary matrix (or a binary image). 
    Since convolution has great advantages in extracting spatial features, the convolutional neural network is an excellent tool in spatial domain. 
    (b) Model work in temporal domain. 
    In temporal domain, the base sequence (represented by the binary matrix) is embedded into a sequance of high-dimensional vector, 
    in which the recurrent neural network perform better.
    In addition, we note that the last layers of neural network are usually full connection structure (not necessarily), 
    which greatly increases the difficulty of understanding the decisions of model.
    }\label{fig:01}
\end{figure}

 where $E$ is the identity matrix, and we define $B$ as the second-order preference matrix, 
 it can explain how a particular pattern containing two nucleotides affects the base sequence, 
 for example, how does the degree of preference for the G adjacent to PAM be affected by a T at 2 bases upstream?
 The previous work (even based on conventional machine learning algorithms with high interpretability) 
 was limited to dimers which can only work on the adjacent base pairs \citep{Liu2019} 
 or have to perform complex feature engineering \citep{MuhammadRafid2020}.
 In light of the above, we believe it is essential to develope a model which can not only match deep learning based model in performance, 
 but also be comparable to conventional machine learning algorithms in interpretability.

@@

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{matrix}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{matrix}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.

Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.

@@

@@

As mentioned above, the predicted value $y$ should satisfy the following equation
y=\sum_i^lW_i\tilde{A}_i\cdot {X_e}_i\label{eq:12}
where $W\in\mathbb{R}^l$ denotes a non-negative weight vector acquired through a full connection layer, $W_i$ denotes the $i$-th dimension of the vector $W$. 
Equation~(\ref{eq:12}) is the same formally as Equation~(\ref{eq:04}). 

@@

\begin{table}[!t]
\processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
head2 & head3 & head4\\\midrule
row1 & row1 & row1 & row1\\
row2 & row2 & row2 & row2\\
row3 & row3 & row3 & row3\\
row4 & row4 & row4 & row4\\\botrule
\end{tabular}}{This is a footnote}
\end{table}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

@@

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text

@@

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text

@@

Text
text text text\vspace*{-12pt}

>>>>

 Specifically, we use a vector $A_i$ to build the first-order original preference at position $i$ within sgRNA sequence, then
 define $\tilde{A}$ as the first-order combine preference matrix (or just first-order preference), which means $\tilde{A}$ can be expressed linearly by $A$ as follow
\begin{equation}
\tilde{A}=BA\label{eq:04}
\end{equation}
where the weight matrix $B\in\mathbb{R}^{l\times l}$ is learned through attention mechanism. 
Then the predicted value can be expressed as
\begin{equation}
y=\sum_i^lW_i\tilde{A}_i\cdot {X_e}_i\label{eq:05}
\end{equation}
 where $W\in\mathbb{R}^l$ denotes a non-negative weight vector, $W_i$ denotes the weight of the $i$-th position, ${X_e}_i$ denotes the embedding vector of the nucleotide at $i$-th position.
 Assume that in Equation~(\ref{eq:04}) the first-order original preference at each position is relatively independent and weight matrix $B$ is independent of any first-order original preference, 
 then the total differential of the first-order preference at $i$-th position $\tilde{A}_i$ could be written as
\begin{equation}
d\tilde{A}_i=\sum_j^lB_{ij}EdA_j\label{eq:07}
\end{equation}
where $E$ is the identity matrix, and we call $B$ as the second-order preference matrix, 
it can explain how a particular pattern containing two nucleotides affects the base sequence, 
for example, how does the degree of preference for the G adjacent to PAM be affected by a T at 2 bases upstream?

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{category.png}}
    \caption{Two categories of the deep learning models used in sgRNA related task. 
    (a) Model work in spatial domain. 
    In spatial domain, the base sequence is encoded into a binary matrix (or a binary image). 
    Since convolution has great advantages in extracting spatial features, the convolutional neural network is an excellent tool in spatial domain. 
    (b) Model work in temporal domain. 
    In temporal domain, the base sequence (represented by the binary matrix) is embedded into a sequance of high-dimensional vector, 
    in which the recurrent neural network perform better.
    In addition, we note that the last layers of neural network are usually full connection structure (not necessarily), 
    which greatly increases the difficulty of understanding the decisions of model.
    }\label{fig:01}
\end{figure}
 The previous work (even based on conventional machine learning algorithms with high interpretability) 
 was limited to dimers which can only work on the adjacent base pairs \citep{Liu2019} 
 or have to perform complex feature engineering \citep{MuhammadRafid2020}.
 In light of the above, we believe it is essential to develope a model which can not only match deep learning based model in performance, 
 but also be comparable to conventional machine learning algorithms in interpretability.

@@

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{matrix}\right.\label{eq:11}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.

@@

Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.

@@

As mentioned above, the predicted value $y$ should satisfy Equation~(\ref{eq:05}). 

@@

@@

Text Text  Text

@@

@@

Text
\begin{table}[!t]
    \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
    head2 & head3 & head4\\\midrule
    row1 & row1 & row1 & row1\\
    row2 & row2 & row2 & row2\\
    row3 & row3 & row3 & row3\\
    row4 & row4 & row4 & row4\\\botrule
    \end{tabular}}{This is a footnote}
\end{table}

\vspace*{-12pt}
